### 2011-01-28 trunk/papers/webist2011/abstract.tex
super-mega shortening done. Down to 6 pages from 10 1/2
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1211 d8b0c59b-6544-0410-94d9-83a5658a058e

B2B systems integration was revolutionised by the introduction of web services,
transforming the way enterprise systems communicate and strengthening the  relationships with providers and customers. The main advantage seen by companies was an increase of operational efficiencies, and a reduction of costs. 
In this scenario, highly qualified software developers are responsible for the integration of services with other systems, through the analysis of formal service descriptions or human-readable specifications. 
However, this model fails when targeting the long tail of enterprise software demand, the end-users. 
Discovery and consumption of web services are difficult tasks for end-users. 
This means that potential long tail of end-users creating task-specific applications from existing services is as of yet completely untapped. 
This paper presents an approach to facilitate the discovery and consumption of business web services by end-users, closing the gap between the two. The approach includes: 
\begin{inparaenum}[(a)]
	\item a method to generate ready-to-use web service wrappers, and 
	\item a catalogue which users can browse to search for web services fitting their needs. 
\end{inparaenum}

>>>>

In B2B systems integration and web services, many companies see the advantage of increased operational efficiencies and a reduction of costs. 
In this scenario, highly qualified software developers are responsible for the integration of services with other systems.
However, this model fails when targeting the long tail of enterprise software demand, the end-users. 
Discovery and consumption of web services are difficult tasks for end-users. 
This means that potential long tail of end-users creating task-specific applications from existing services is as of yet completely untapped. 
This paper presents an approach to facilitate the discovery and consumption of business web services by end-users, closing the gap between the two. The approach includes: 
\begin{inparaenum}[(a)]
	\item a catalogue which users can browse to search for web services fitting their needs, and
	\item a method to generate ready-to-use web service wrappers to use in the catalogue.
\end{inparaenum}

### 2011-01-28 trunk/papers/webist2011/paper_sec_introduction.tex
super-mega shortening done. Down to 6 pages from 10 1/2
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1211 d8b0c59b-6544-0410-94d9-83a5658a058e

Business-to-business (B2B) integration is still a significant challenge, often requiring extensive efforts in terms of different aspects and technologies for protocols, architectures or security. Solutions based on open standards 
such as RosettaNet, ebXML, the Web Service Description Language (WSDL), Universal Description, Discovery and Integration (UDDI) or the Simple Object Access Protocol (SOAP)
have reduced the complexity of integrating different business applications between different companies and partners. 
% In this sense, Web services and XML standards such as RosettaNet, ebXML or OAG have been a boon to the world of B2B. By web services standards we are referring to the following open standards: Web Services Description Language (WSDL --- to describe), Universal Description, Discovery and Integration (UDDI --- to advertise and syndicate), Simple Object Access Protocol (SOAP --- to communicate) and Web Services Flow Language (WSFL --- to define work flows). 
% A common scenario  in their usage is a company using WSDL to describe its web services, publishing them through a repository using UDDI, while these web services use SOAP-based messages to achieve dynamic integration between different, disparate applications.

While the adoption of these standards has offered some advantages in business-to-consumer (B2C) integration as well,  the interaction and consumption of web services still requires programming skills and a deep understanding of the  technology, which poses an obstacle for an end-user with limited knowledge on the matter.% (in the sense as defined in~\cite{fuchsloch2010}). 
Regarding the discovery of the right service for the right task --- a crucial requirement for the dynamic use of WS --- most publishing platforms are syntax-based, %leading to poor precision and recall in finding the correct service, 
making it difficult to navigate through a large number of web services~\cite{pilioura_acm2009}, 
therefore preventing end-users from performing these tasks. 
Solutions such as Semantic Web Services (SWS) promised many advantages in discovery, composition and consumption of web services, independently of the provider's platform, location, service implementation or data format. 
However, as of yet they have not been widely adopted, 
possibly because the perceived potential benefits of SWS did not justify the additional investments required for businesses~\cite{shi2007}.
In a similar vein as B2C, governments are currently opening their data to the public as linked data. 
So far, the efforts are restricted to data only, so that users are left alone in finding applications of that data. 
A natural further development in the public sector would be that their services for administration tasks and single-window applications will be opened as well. Hence, governments will benefit from the research done around these concepts and technologies.

The motivation of our work has been strongly influenced by the end-users' needs. We are targeting users which are non-skilled in term of programming and software development, empowering them with a platform to discover and consume web services in a straight-forward manner.
The web services and their specifications are not published right away in the platform, since this would simply mean to re-implement UDDI.
However, their definition and behaviour are wrapped in what we call a ``web service wrapper''. The products resulting from the wrapping process are two artifacts: a specific definition of the web service to be used by this tool, and a ready-to-consume (inside a web browser) piece of code with the proper functions to invoke the web service.

>>>>

Business-to-business (B2B) integration is still a significant challenge, often requiring extensive efforts in terms of different aspects and technologies for protocols, architectures or security.
While the adoption of open standards 
such as RosettaNet, ebXML, the Web Service Description Language (WSDL), Universal Description, Discovery and Integration (UDDI) or the Simple Object Access Protocol (SOAP)
have reduced the complexity of integrating business applications between different companies and partners,
% In this sense, Web services and XML standards such as RosettaNet, ebXML or OAG have been a boon to the world of B2B. By web services standards we are referring to the following open standards: Web Services Description Language (WSDL --- to describe), Universal Description, Discovery and Integration (UDDI --- to advertise and syndicate), Simple Object Access Protocol (SOAP --- to communicate) and Web Services Flow Language (WSFL --- to define work flows). 
% A common scenario  in their usage is a company using WSDL to describe its web services, publishing them through a repository using UDDI, while these web services use SOAP-based messages to achieve dynamic integration between different, disparate applications.
and offered some advantages in business-to-consumer (B2C) integration as well, the interaction and consumption of web services still requires programming skills and a deep understanding of the  technology, which poses an obstacle for an end-user with limited knowledge of the matter. % (in the sense as defined in~\cite{fuchsloch2010}). 
Regarding the selection of the right service for the right task --- a crucial requirement for the dynamic use of web services --- most publishing platforms are syntax-based, %leading to poor precision and recall in finding the correct service, 
making it difficult to navigate through a large number of web services~\cite{pilioura_acm2009}, 
preventing end-users from performing these tasks. 
Solutions such as Semantic Web Services (SWS) promised many advantages in this respect.
% discovery, composition and consumption of web services, independently of the provider's platform, location, service implementation or data format. 
However, as of yet they have not been widely adopted, 
possibly because the perceived potential benefits did not justify the additional investments~\cite{shi2007}.
% In a similar vein as B2C, governments are currently opening their data to the public as linked data. 
% So far, the efforts are restricted to data only, so that users are left alone in finding applications of that data. 
% A natural further development in the public sector would be that their services for administration tasks and single-window applications will be opened as well. Hence, governments will benefit from the research done around these concepts and technologies.

The motivation of our work has been strongly influenced by the end-users' needs. We are targeting users which are non-skilled in programming and software development, empowering them with a platform to select and consume web services in a straight-forward manner.
Rather than publishing services directly in our platform, we leave existing third-party services untouched and instead integrate them through service wrappers. The products resulting from the wrapping process are two artifacts: a specific definition of the web service to be used by this tool, and a ready-to-use piece of code with the proper functions to invoke the web service.

### 2011-01-28 trunk/papers/webist2011/paper_sec_related_work.tex
super-mega shortening done. Down to 6 pages from 10 1/2
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1211 d8b0c59b-6544-0410-94d9-83a5658a058e

Web services have been around for a long time since they first appeared.
One of the most important claims about their benefits has been (syntactic) interoperability between third-party systems and applications based on different platforms / programming languages. 
System integration, within a company or between systems from different enterprises, became easier with the adoption of web service standard technologies, in particular the Web Service Description Language (WSDL), the most extensive standard used for the definition of web services. 
Many integrated development environments (IDEs) can deal with WSDL to facilitate the integration task, however it is still required that developers read and understand their specifications, and implement part of the logic programmatically. 
As a step forward, there are several tools which facilitate the interaction with data sources and services on the Web. Yahoo! Pipes, Apatar, JackBe Presto, Microsoft Popfly (now defunct) and NetVibes, among others provide a set of modules to access different kind of data sources, such as RSS feeds, a given web page (HTML code), Flickr images, Google base or the Yahoo! search engine, databases (MySQL, PostgreSql, Oracle), and powerful enterprise systems such as Salesforce CRM, SugarCRM and Goldmine CRM. 
However, none of the solutions in the market really facilitates end-users to build  their own applications or widgets allowing the interaction with web services created by third-party providers. The solutions found are mainly data-oriented (RSS feeds, databases, raw text). Several tools permits some sort of (web) service integration, but are meant to be used within an enterprise level by savvy business users or developers. The solution presented in this paper leverage the possibility of integrating REST or SOAP-based web service inside visual applications running in a browser (e.g. widgets) by providing a platform to create, publish and discover web services wrappers.

In the context of publishing and discovering Web services, service providers have well-known and widely used technologies to accomplish the task of publication, such as the Universal Description, Discovery and Integration (UDDI) or WS-Discovery. The UDDI service registry specification~\cite{uddi2004} is currently one of the core standards in the Web service technology stack and an integral part of every major SOA vendor's technology strategy and offers to requesters the ability to discover services via the Internet. 
In short, UDDI serves as a centralised repository of WSDL documents. A similar concept is iServe~\cite{pedrinaci_ores2010}. This platform aims to publish web services as what they called Linked Services --- linked data describing services ---, storing web service definitions as semantic annotations, so that other semantic web-aware applications may take advantage of it. However, the platform does not deal with the step from the definition to the consumption of the services.

>>>>

Web services have been around for a long time.
One of the most important claims about their benefits has been (syntactic) interoperability between third-party systems and applications based on different platforms / programming languages. 
System integration, within a company or between systems from different enterprises, became easier with the adoption of web service standard technologies such as WSDL. 
While many integrated development environments (IDEs) can deal with WSDL to facilitate the integration task, trained developers are still required for this.
% however it is still required that developers read and understand their specifications, and implement part of the logic programmatically. 
As a step forward, there are several tools which facilitate the interaction with data sources and services on the Web. Yahoo! Pipes, Apatar, JackBe Presto and NetVibes, among others provide a set of modules to access different kind of data sources, such as RSS feeds, a given web page (HTML code), Flickr images, databases, and even powerful enterprise systems such as Salesforce CRM or Goldmine CRM. 
However, none of these solutions facilitate end-users to build their own applications allowing the interaction with web services created by third-party providers. The solutions found are mainly data-oriented (RSS feeds, databases, raw text). Several tools permits some sort of (web) service integration, but are meant to be used within an enterprise level by savvy business users or developers. The solution presented in this paper leverages the possibility of integrating RESTful or SOAP-based web services inside browser-based applications (i.e., widgets or gadgets) by providing a platform to create, publish and select web services wrappers.

In the context of publishing and discovering Web services, service providers have well-known and widely used technologies to accomplish the task of publication, such as the Universal Description, Discovery and Integration (UDDI)~\cite{uddi2004}. UDDI serves as a centralised repository of WSDL documents. A similar concept is iServe~\cite{pedrinaci_ores2010}. This platform aims to publish web services as what they called Linked Services --- linked data describing services ---, storing web service definitions as semantic annotations, so that other semantically aware applications may take advantage of it. However, the platform does not handle the step from definition to consumption of the services.

### 2011-01-28 trunk/papers/webist2011/paper_sec_use_case.tex
super-mega shortening done. Down to 6 pages from 10 1/2
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1211 d8b0c59b-6544-0410-94d9-83a5658a058e

The intention of this paper is to demonstrate the advances made regarding web service discovery and consumption through two artifacts developed as part of our research: the \emph{publishing and discovery platform} and the \emph{service wrapper tool}. While these artifacts may be deployed and used separately by third-party applications, they were originally developed to form the backbone of the FAST platform~\cite{hoyer2009fast}.
FAST constitutes a novel approach to application composition and business process definition from a user-centric perspective. It is aimed at allowing users without previous programming experience to create their own situational applications by visually combining different \emph{building blocks}, such as graphical forms and back-end services, based on their inputs and outputs (or \emph{pre-} and \emph{post-conditions}). The work covered in this paper are the components highlighted and surrounded by a dashed frame in the Fig.~\ref{fig:fast_architecture}. Communication within the platform is mostly done via a RESTful API, using JSON as an exchange syntax.
%These components have public interfaces to communicate with any application developed by a third-party developer or company, and as described in Sects. 3 and 4, are fully integrated and present a high cohesion with the rest of the components of the FAST architecture.

@@

In order to facilitate a better understanding of these components, we describe a number of concepts related to FAST in this section. 
A \emph{gadget} is the end product of the platform, ready to be run in any ordinary web browser, usually through a mashup platform. 
Within FAST, the platform-independent gadget is called \emph{screenflow}, which comprises a set of domain-specific \emph{screens} connected through their \emph{pre-} and \emph{post-conditions}. A screen is the most complex building block fully functional by itself. 
Visually, it is similar to a tab in a tabbed application.
It is composed by a \emph{form} conveying the graphical interface, and a set of \emph{operators} and \emph{back-end services}, wrapped into so-called \emph{service resources}.

The service wrapper tool is used to create wrappers for third-party web services, transforming them into building blocks to be stored and reused within the FAST platform. The publishing platform presented, also called \emph{catalogue} within the FAST platform, covers several important purposes, such as storage, indexing, publication and search of gadgets, gadget building blocks and user profiles, as well as some level of ontology mediation to facilitate the integration of services from different parties (first steps in this direction are outlined in \cite{Ambrus:2009it}).

>>>>

The intention of this paper is to demonstrate the advances made regarding web service discovery and consumption through two artifacts developed as part of our research: the \emph{publishing and discovery platform} and the \emph{service wrapper tool}. While these artifacts may be deployed and used separately by third-party applications, they were originally developed to form the backbone of the FAST platform~\cite{hoyer2009fast}.
FAST constitutes a novel approach to application composition from a user-centric perspective. It is aimed at allowing users without previous programming experience to create their own situational applications by visually combining different \emph{building blocks}, such as graphical forms and back-end services, based on their inputs and outputs (or \emph{pre-} and \emph{post-conditions}). The work covered in this paper are the components highlighted in Fig.~\ref{fig:fast_architecture} by a dashed line. Communication within the platform is mostly done via a RESTful API, using JSON as an exchange syntax.
%These components have public interfaces to communicate with any application developed by a third-party developer or company, and as described in Sects. 3 and 4, are fully integrated and present a high cohesion with the rest of the components of the FAST architecture.

@@

In order to provide a better understanding of these components, we describe a number of concepts related to FAST in this section. 
A \emph{gadget} is the end product of the platform, ready to be run in any ordinary web browser, usually through a mashup platform. 
An undeployed gadget is also called \emph{screenflow}, which comprises a set of \emph{screens} connected through their \emph{pre-} and \emph{post-conditions}. A screen is the most complex building block fully functional by itself,
visually similar to a tab in a tabbed application.
It is composed by a \emph{form} conveying the graphical interface, and a set of \emph{operators} and \emph{back-end services}, wrapped into so-called \emph{service resources}.

The publishing platform, also called \emph{catalogue}, covers several important purposes, such as storage, indexing, publication and search of gadgets, gadget building blocks and user profiles.
% , as well as some level of ontology mediation to facilitate the integration of services from different parties (first steps in this direction are outlined in \cite{Ambrus:2009it}).
The service wrapper tool is used to create wrappers for third-party web services, transforming them into building blocks to be stored and reused within the FAST platform.

### 2011-01-28 trunk/papers/webist2011/paper_sec_wrapping.tex
super-mega shortening done. Down to 6 pages from 10 1/2
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1211 d8b0c59b-6544-0410-94d9-83a5658a058e

In our approach, wrapping web services is done in two steps: a first step is in charge of the construction of a service request and a second step analyses the response received from the execution of the service, allowing the analysis of response data and the mapping to domain-specific concepts. From this input, we generate JavaScript code to be embedded into web gadgets. This JavaScript code takes query parameters, constructs the service request, analyses the service response and returns result data in the desired format.

@@

In this paper we illustrate the interaction with services on simple GET requests as supported e.g. by REST services. Support for other services like POST based services is under construction. In case of a simple GET request, a service request is assembled using a certain URL and a set of parameters. (A POST request would have a text block to be posted. In addition, other header data may be needed to transfer e.g. session ids). As an example, the Ebay Shopping web service will be studied. As documented on the corresponding web site, the following URL is invoked to retrieve a list of items corresponding to the search keywords \emph{USB}:

\begin{listing}
\begin{verbatim}
http://open.api.sandbox.ebay.com/shopping?
  appid=KasselUn-efea-4b93-9505-5dc2ef1ceecd&
  version=517&callname=FindItems&
  ItemSort=EndTime&QueryKeywords=USB&
  responseencoding=XML
\end{verbatim}
\end{listing}

As you may see, the URL invoked is \url{http://open.api.sandbox.ebay.com/shopping} and the parameters used in the example are:
\begin{description}
	\item[appid] this is the application ID obtained to use the API.
	\item[version] the API version.
	\item[callname] in this case \emph{FindItems} to search through all items in Ebay.
	\item[ItemSort] sorting method for the list of items.
	\item[QueryKeywords] list of keywords.
	\item[responseencoding] format of the response message obtained by the invocation of the request.
\end{description}

The above URL for searching items in Ebay is followed by the query parameters, which take the form \textit{argument=value}. In this example, the only relevant parameter the end user may want to provide is \emph{querykeywords}. Thus the end user may use some input field to provide a query keyword and pass this value to our service wrapper operation. Other parameters can be set to a default value. However, in order to develop a more generic wrapper to a service, more or even all parameters might be passed by the user.

To define the desired input parameters and the type of the expected result of a new service wrapper the service wrapper tool provides a web-based form that allows the editing of these entries, see Fig.~\ref{fig:construct_pre_post_conditions}. Note, we allow to edit example values for the input parameters. These example values may be used to test the service wrapper.

\begin{figure}
  \begin{center} \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithPortDefinitions.png}
    \caption{Configuring parameters of a service wrapper}
    \label{fig:construct_pre_post_conditions}
  \end{center}
\end{figure}

@@

An example of service request construction can be found in the Figure~\ref{fig:construct_service_request}. In the top input field the user may drop an example HTTP request, taken e.g. from the service's documentation\footnote{\url{http://developer.ebay.com/products/shopping/}}. The tool analyses the example request and in the middle of the screen a form for editing the request parameters is provided. In our example, the user has connected the \textit{QueryKeywords} parameter with the input parameter \textit{search\_key} by adding a corresponding reference to the value field of that parameter. In addition, we have retrieved an access key which has been entered as value for the \textit{appid} parameter.

\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithRequestExample.png}
    \caption{Constructing the service request URL and its parameters}
    \label{fig:construct_service_request}
  \end{center}
\end{figure}

Below the request parameter editing form of Figure~\ref{fig:construct_service_request}, a \textit{Send Request} button allows to validate the constructed URL by sending it to the specified service address (via a server relay). 
The service's response is shown on the bottom of the page. This gives the user a fast feedback whether the constructed request works as desired.

\subsubsection{Limitations} % (fold)
\label{ssub:limitations}

It is worth pointing out that currently the wrapper tool is able to construct input ports for the wrappers using just basic types. To construct URL requests from more complex input parameters, as e.g. a person object, we are currently developing access operations that will allow to access the values of fields of complex objects. For example, the expression \emph{customer.address} may refer to the \emph{address} field of a \emph{customer} parameter.

% subsubsection limitations (end)

@@

A web service can choose to send a response message in any desired structured format. In our example, we assume an XML tree, but a JSON structure is equally possible.

\subsubsection{Translation of the Response Tree into Facts} % (fold)
\label{ssub:translation_xml_into_facts}

\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithTransformationRules.png}
    \caption{Interactive, rule based transforming of an XML response to domain objects}
    \label{fig:response_service_execution}
  \end{center}
\end{figure}

Once the service response has been retrieved, the transformation tab of the wrapper tool shows it as an interactive object tree, as seen on the left side of Fig.~\ref{fig:response_service_execution}.

A transformation rule is used to analyse the response data and to generate domain-specific objects from concepts from the ontologies used by the parameters of the different building blocks. A transformation rule is composed of three elements, as seen in the middle part of Fig.~\ref{fig:response_service_execution}. First, the \textit{from} field indicates the source elements to be translated by the rule. Second, the type of the rule will be set, taking one of the following values: \emph{createObject}, \emph{fillAttributes} or \emph{dummy}. And third, the target of the rule specifies a certain concept or attribute, to be created or filled. A detailed explanation of the type of actions to be trigger from the transformation rules is as follows:

@@

specifies the creation of a new domain object. The type of that new object is provided in the third compartment. In the example being explained, the root rule searches for XML elements with tag name \emph{FindItemsResponse} and for each such element a \emph{List} object is created. The resulting objects are shown in a facts tree in the right of Figure ~\ref{fig:response_service_execution}.

@@

does not create a new object but it fills the value of the attribute provided as third part of such rules. In our example, the third transformation rule searches for XML elements with tag name \emph{Title}. Note, the rule is a sub-rule of the second rule, which generates \emph{Product} objects. Thus, the sub-rule searches for \emph{Title} tags only in the subtree of the XML data that has been identified by an application of the parent rule before. For example, the \emph{Item} rule may just have been applied to the first \emph{Item} element of the XML data. Then, the \emph{Title} rule is applied only to the first \emph{Item} sub-tree of the XML data and thus it will find only one \emph{Title} element in that sub-tree (not visible in Figure ~\ref{fig:response_service_execution} ). The value of that \emph{Title} element is then transfered to the \emph{productName} attribute of the corresponding \emph{Product} object.

@@

does not create or modify any objects but such rules are just used to narrow the search space for their sub-rules. For example, in Amazon product data, the XML data for an item contains sections for \emph{minimum price}, \emph{maximum price}, and \emph{average price}. Each such section contains the \emph{plain price} and the \emph{formatted price}. Thus, in the Amazon case, a rule that searches for \emph{formatted price} elements within an \emph{Item} element would retrieve three matches. Using a dummy rule, we may first search for \emph{minimum price} elements and then search for \emph{formatted price} elements within that sub-tree.

@@

Our tool follows an interactive paradigm. Therefore, any time a change to a transformation rule is done, the transformation process is triggered and the resulting facts tree is directly shown. This process helps the user to deal with the slightly complex semantics of the transformation rules avoiding errors or mistakes. In addition, our tool is ontology-driven, therefore the service wrapper designer retrieves the domain-specific types from a backend server (such as the catalogue discussed in Section~\ref{sec:discovery}) together with the structure of each type, i.e. together with a description of the attributes of each object. Thus, the transformation rule editor is able to provide selection boxes for the target element of the rules. For a \emph{createObject} rule, this selection box shows the object types available for that domain. For the \emph{fillAttributes} rules, the selection box shows the attributes of the object type chosen in the parent rule. 
% In addition, we may provide some analysis tool, which will help to guarantee that the objects generated by the transformation rules conform to the object types defined in the corresponding conceptual model (see Section~\ref{ssec:ontology}). This helps to ensure that the objects generated by the designed service wrapper will be compatible for input parameters of subsequent filter steps and or gadgets.

@@

\subsection{Generating a Resource Adapter} % (fold)
\label{sub:generating_a_resource_adapter}

Once the wrapping of a service has been defined and tested in the service wrapper tool, we generate an implementation of the desired Resource Adapter in XML, HTML, and JavaScript, ready to be deployed and executed inside a web gadget.

@@

\subsection{Limitations} % (fold)
\label{sub:limitations}

The rule driven approach presented above is somewhat limited. It is deliberately restricted to such a simple rule mechanism in order to keep things simple enough for end-users. Still, the selected approach suffices for many practical and real world cases. As a more complex example, the XML data for a person may provide two different tags for the first and the last name of a person. Contrarily, a person fact which conforms to a certain ontology for that domain may provide only one \emph{fullname} attribute that shall be filled by a concatenation of the first and the last name. To achieve this, the \textit{from} field of that transformation rule might look like: \texttt{lastname"', "`firstname}. We are also able to do some navigation in the XML tree to follow XRef elements. For example the attribute \texttt{grandmother} could be filled using \texttt{mother.mother} in the \textit{from} field.

However, there are some transformations that these rules cannot perform. For example, we do not support any mathematical operations. Thus, transforming e.g. Fahrenheit into Celsius temperatures is not supported. To cover such  cases, intermediate object formats can be used which would allow generating objects to be further processed by additional filters. Such additional filters may be realised using (hand coded) operators, since some generic operators can act as filters for aggregation and conversions of objects from multiple sources. Then, service wrappers in combination with these filter operators will allow covering these complex cases.

% subsection limitations (end)

>>>>

Before third-party services can be used by our platform, they need to be provided with a service wrapper. In our approach, this is done in two steps: 
\begin{inparaenum}[(i)]
	\item constructing an exemplary service request, and 
	\item analysing the response received from the execution of the service, allowing the mapping of the response data to domain-specific concepts to be used as pre- and post-conditions.
\end{inparaenum}
From this input, the tool then generates the actual wrapper --- JavaScript code to be embedded into web gadgets.

@@

In this section we illustrate the interaction with RESTful services on simple GET requests\footnote{Other methods such as POST or SOAP-based services are also supported.}. 
In this case, a service request is assembled using its URL and a set of parameters. As an example, we will look at the eBay Shopping web service. 
To define the desired pre-condition and the type of post-condition of a new service wrapper, the service wrapper tool provides a form (not shown here). For the current example, we assume that the user has defined a pre-condition \emph{search\_key}.

We are invoking the service to retrieve a list of items corresponding to the search keyword ``USB'':

\begin{lstlisting}
http://open.api.sandbox.ebay.com/shopping?
  appid=KasselUn-efea-xxx&version=517&
  callname=FindItems&ItemSort=EndTime&
  QueryKeywords=USB&ResponseEncoding=XML
\end{lstlisting}

% The parameters used in the example are:
% \textbf{appid} (the application ID obtained to use the API),
% \textbf{version} (the API version),
% \textbf{callname} (in this case \emph{FindItems} to search through all items in eBay),
% \textbf{ItemSort} (sorting method for the list of items),
% \textbf{QueryKeywords} (list of keywords), and
% \textbf{ResponseEncoding} (format of the response message obtained by the invocation of the request).

% Note that we allow to edit example values for the input parameters, which can be used to test the service wrapper.

% \begin{figure}
%   \begin{center} \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithPortDefinitions.png}
%     \caption{Configuring parameters of a service wrapper}
%     \label{fig:construct_pre_post_conditions}
%   \end{center}
% \end{figure}

@@

The various parameters in the example are defined by the service provider. However, the most relevant one for our example is \emph{QueryKeywords}, which communicates to the services what we are looking for. 
An example of service request construction is shown in the Fig.~\ref{fig:construct_service_request}. In the ``Request'' input field the user enters an example HTTP request, e.g. the one above. The tool analyses the request and provides a form for editing the request parameters. In our example, the user has connected the service's \textit{QueryKeywords} parameter with the wrapper's \textit{search\_key} pre-condition. Requests constructed in this way can then be sent off to the service, whose response is shown in the bottom text area.

\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithRequestExample.png}
    \caption{Constructing the service request}
    \label{fig:construct_service_request}
  \end{center}
\end{figure}

% \subsubsection{Limitations} % (fold)
% \label{ssub:limitations}
% 
% It is worth pointing out that currently the wrapper tool is able to construct input ports for the wrappers using just basic types. To construct URL requests from more complex input parameters, as e.g. a person object, we are currently developing access operations that will allow to access the values of fields of complex objects. For example, the expression \emph{customer.address} may refer to the \emph{address} field of a \emph{customer} parameter.  
% 
% % subsubsection limitations (end)

@@

% A web service can choose to send a response message in any desired structured format. In our example, we assume an XML tree, but a JSON structure is equally possible.

% \subsubsection{Translation of the Response Tree into Facts} % (fold)
% \label{ssub:translation_xml_into_facts}

Once the service response has been retrieved, the transformation tab of the wrapper tool shows it as an interactive object tree, as seen in Fig.~\ref{fig:response_service_execution}.

\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithTransformationRules.png}
    \caption{Rule-based transformation of service response}
    \label{fig:response_service_execution}
  \end{center}
\end{figure}

A transformation rule is used to analyse the response data and generate the wrapper's post-conditions. Such a rule is composed of three elements, as seen in the middle part of Fig.~\ref{fig:response_service_execution}. The \textit{from} field indicates the source elements to be translated by the rule. The second part defines the type of rule (see below). Thirdly, the target of the rule specifies a certain concept or attribute, to be created or filled. Rules can be chained together in a rule tree. A detailed explanation of the different rule types is as follows:

@@

specifies the creation of a new post-condition of the type specified in the third part of the rule. In the example in Fig.~\ref{fig:construct_service_request}, the root rule could search for source elements with name \emph{FindItemsResponse} and create a post-condition of type \emph{List} for each such element. The resulting objects are shown in a facts tree in the right of the figure.

@@

only ever appear as sub-rules in the rule tree. They match source elements defined in the first part of the rule. The third part defines the attribute of the super-rule that will be filled (e.g., \emph{fullName} in Fig.~\ref{fig:response_service_execution}).
% In Fig.~\ref{fig:response_service_execution}, the third transformation rule matches elements with the name \emph{Title}. The 
% Note that this rule is a sub-rule of the second rule, which generates \emph{Product} objects. Thus, the sub-rule searches for \emph{Title} tags only in the subtree of the XML data that has been identified by an application of the parent rule before. For example, the \emph{Item} rule may just have been applied to the first \emph{Item} element of the XML data. Then, the \emph{Title} rule is applied only to the first \emph{Item} sub-tree of the XML data and thus it will find only one \emph{Title} element in that sub-tree (not visible in Figure ~\ref{fig:response_service_execution} ). The value of that \emph{Title} element is then transfered to the \emph{productName} attribute of the corresponding \emph{Product} object.

@@

does not create or modify any objects but instead narrows the search space for their sub-rules, by selecting certain elements in the source tree and ignoring others.

% E.g., in Amazon product data, each item contains elements for \emph{minimum price}, \emph{maximum price}, and \emph{average price}. Each such element in turn contains a \emph{plain price} and \emph{formatted price}. In this case, a rule that searches for \emph{formatted price} elements within an \emph{Item} element would retrieve three matches. Using a dummy rule, we can first search for \emph{minimum price} elements and then search for \emph{formatted price} elements within that sub-tree.

@@

Our tool follows an interactive paradigm. Any time a change to a transformation rule is done, the transformation process is triggered and the resulting facts tree is shown. This aids the user to deal with the complexity of the transformation rules, preventing errors or mistakes. In addition, the tool is ontology-driven, meaning that possible post-condition types and their attributes are selected from domain ontologies loaded in the system.
% therefore the service wrapper designer retrieves the domain-specific types from a backend server (such as the catalogue discussed in Section~\ref{sec:discovery}) together with the structure of each type, i.e. together with a description of the attributes of each object. Thus, the transformation rule editor is able to provide selection boxes for the target element of the rules. For a \emph{createObject} rule, this selection box shows the object types available for that domain. For the \emph{fillAttributes} rules, the selection box shows the attributes of the object type chosen in the parent rule. 
% In addition, we may provide some analysis tool, which will help to guarantee that the objects generated by the transformation rules conform to the object types defined in the corresponding conceptual model (see Section~\ref{ssec:ontology}). This helps to ensure that the objects generated by the designed service wrapper will be compatible for input parameters of subsequent filter steps and or gadgets.

@@

\subsection{Generating the Service Wrapper} % (fold)
\label{sub:generating_a_resource_adapter}

Once the wrapping of a service has been defined and tested in the tool, we generate an implementation of the desired specification in XML, HTML, and JavaScript, ready to be deployed and executed inside a web gadget.

@@

% \subsection{Limitations} % (fold)
% \label{sub:limitations}
% 
% The rule driven approach presented above is somewhat limited. It is deliberately restricted to such a simple rule mechanism in order to keep things simple enough for end-users. Still, the selected approach suffices for many practical and real world cases. As a more complex example, the XML data for a person may provide two different tags for the first and the last name of a person. Contrarily, a person fact which conforms to a certain ontology for that domain may provide only one \emph{fullname} attribute that shall be filled by a concatenation of the first and the last name. To achieve this, the \textit{from} field of that transformation rule might look like: \texttt{lastname"', "`firstname}. We are also able to do some navigation in the XML tree to follow XRef elements. For example the attribute \texttt{grandmother} could be filled using \texttt{mother.mother} in the \textit{from} field. 
% 
% However, there are some transformations that these rules cannot perform. For example, we do not support any mathematical operations. Thus, transforming e.g. Fahrenheit into Celsius temperatures is not supported. To cover such  cases, intermediate object formats can be used which would allow generating objects to be further processed by additional filters. Such additional filters may be realised using (hand coded) operators, since some generic operators can act as filters for aggregation and conversions of objects from multiple sources. Then, service wrappers in combination with these filter operators will allow covering these complex cases.
% 
% subsection limitations (end)
